Hadoop Image Processing Framework

### 2015-03-15 Paper.tex
Nearly finished with all editing except for related work.

\section{Methodology}
The Hadoop Image Processing Framework is intended to provide users
with an accessible, easy-to-use tool for developing large-scale image
processing applications.

@@

\section{Image Processing Computations}
To explore the Hadoop Image Processing Framework's capabilities and
performance, we implemented multiple variations of existing
widely-used image processing algorithms:
\begin{itemize}
\item A stand-alone implementation running on a single node with no
  distributed processing capability
\item A generic Hadoop implementation which requires a high level of
  software engineering expertise within the Hadoop framework
\item An implementation that employs the Hadoop Image Processing
  Framework
\end{itemize}

@@

We describe two image processing algorithms using Hadoop Image Processing framework. These image processing algorithms are relatively simple to implement on a single node, implementing these algorithms are complex and inefficient on existing platforms, yet simple to implement using the framework.In the following section we provided image processing computations with the data set described in Section VA. The hardware and software specifications are provided in Section VB as well as observations in Section VC

@@

\subsection{Characteristics of Image Dataset, Hardware and Software}
We computed image processing computations on a dataset which is about 1TB size. The data set is made from the flickr search based on a keyword flight. The data set is extracted using the a FlickrRipper in the framework which implements Ripper Interface. The Ripper interface can be implemented for any custom Rippers to get the image URLs. The example data set is composed of 220000 images at 4.76 MB/image equals to 1 TB.

\subsection{Computer Hardware and Software Characterstics}
We ran the image processing computations on a cluster with 6 nodes and on a desktop computer. Table summarizes the cluster and desktop hardware and software specifications. The cluster nodes differ in terms of CPU speed and RAM. We installed hadoop, cloudera CDH5 and java 1.7 on the cluster to support java code execution. The desktop computer had similar software configuration to the cluster.

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		& \multicolumn{1}{c|}{\textbf{Specs}}                            & \multicolumn{1}{c|}{\textbf{Cluster}}                                                                                                           & \multicolumn{1}{c|}{\textbf{Desktop}}                                                                               \\ \hline
		\multirow{2}{*}{Hardware} & Cluster Nodes                                                  & \begin{tabular}[c]{@{}l@{}}6 datanodes having \\ from 2 to 4 virtual \\ processors with 4 GB\\ RAM and namenode \\ having 8 GB RAM\end{tabular} & \begin{tabular}[c]{@{}l@{}}Intel Xeon @ 2Ghz\\ 8 cores, 16GB of RAM \\ and hyper threading\\ activated\end{tabular} \\ \cline{2-4} 
		& Networking                                                     & 1Gbit/second                                                                                                                                    &                                                                                                                     \\ \hline
		\multirow{3}{*}{Software} & \begin{tabular}[c]{@{}l@{}}Java Virtual\\ Machine\end{tabular} & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                 \\ \cline{2-4} 
		& Hadoop                                                         & \begin{tabular}[c]{@{}l@{}}hadoop 2.5\\ cloudera CDH5\end{tabular}                                                                              &                                                                                                                     \\ \cline{2-4} 
		& \begin{tabular}[c]{@{}l@{}}Operating\\ System\end{tabular}     & CentOS 6.5                                                                                                                                      & CentOS 6.5                                                                                                          \\ \hline
	\end{tabular}
\end{table}

The desktop implementation of the Algorithm is a java program for processing a small image dataset. As soon as the data set is processed the program halts. The desktop implementation starts as a single thread. One image is processed every time. As soon as the image is processed, it starts processing a new image. When the data set size is low, the program runs smooth with best performance. As data set size increases to Tera-bytes the system can't handle such huge amounts of data. To handle such large amounts of data we use the distributed computing using hadoop.

The Algorithm is implemented on a hadoop cluster without using framework. It overcomes the problem of handling large datasets but suffers with a problem of code complexity. The user needs to write a huge amount of code and understand the technical details of hadoop. Writing such huge amounts of code including the custom InputFormats and RecordReaders is cumbersome. Fig.\ref{fig:module-chart} compares the actual lines of code user needs to write to run an Image Processing application in hadoop cluster which makes an average user life harder to use the hadoop features.

The Algorithm is finally implemented on a hadoop cluster using the framework. It overcomes many problems handling large datasets, code complexity, interoperability between image datatypes etc., The process of creating CustomInputFormat and RecordReader is automated as per the input. User can tune the settings using different getter and setter method to get the desired output. The framework is so transparent that the user needs to input few lines of code to process a large image data set. It hides all the MapReduce Technical details and runs the code highly parallelized. The key differences among the three platform-specific implementation lie in computation elasticity and code complexity associated with each platform.

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		\multicolumn{1}{|c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computa-\\ tional\\ Platform\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computational \\ Elasticity\end{tabular}}}      & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Code \\ Complexity\end{tabular}}}                                                                                                                                                  & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Data \\ Locality\end{tabular}}}                                                                           \\ \hline
		Desktop                                                                                             & \begin{tabular}[c]{@{}l@{}}Low : limited by \\ RAM and CPU of \\ the executing\\ computer\end{tabular} & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing the \\ algorithm\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}All data are \\ on local disk\end{tabular}                                                                                            \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ without \\ using \\ framework\end{tabular}                     & \begin{tabular}[c]{@{}l@{}}High: Nodes can\\  be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Heavy : Complexity \\ lies inwriting every \\ module in framework \\ including custom \\ Input Formats and \\ RecordReaders.User \\ should have high \\ technical knowledge on \\ hadoop working.\end{tabular} & \begin{tabular}[c]{@{}l@{}}All data resides\\ in HDFS. Custom\\ InputFormats \\ needsto be \\ created to launch\\ computations \\ where thedata are\end{tabular} \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ using \\ framework\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}High: Nodes can \\ be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing \\ algorithm and is \\ equivalent to writing \\ in Desktop platform\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}All data resides \\ in HDFS. Highly \\ parallelized.\end{tabular}                                                                     \\ \hline
	\end{tabular}
\end{table}

@@

We have documented the following samples in the space of image processing
computations, and hardware and software configurations applied to the selected
data set.
\begin{itemize}
	\item Three image processing computations described in Table 1.
	\item Sampled data set of different image counts.
	\item Hardware cluster configurations used for computations. 
	\item Coding complexities in different environments for the three image processing computations
\end{itemize}

The numerical results are shown in Figure 1 through 6.

Figure \ref{fig:perf-chart} illustrates the performance comparison of Single Node, Hadoop using framework and Hadoop without using the framework for Canny Edge Detection.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{perf-graph}
	\caption{Canny Edge Detection computations executed on single node and on a Hadoop cluster with and without using framework}
	\label{fig:perf-chart}
\end{figure}
It can be observed that the performance hadoop cluster is almost the same with and without using framework. The framework performs well sometimes because of multi-threading is implemented few tasks. Considering the single node performance it can be seen that the single node outperforms hadoop if the input is small as the input size increases handling huge data becomes infeasible. The same pattern is observed with any image processing algorithm.

Figure \ref{fig:module-chart} compares coding complexities of different modules the framework described. The Downloader module and Extractor module are simple to use and require less number of lines of code to configure. The processor module includes the algorithm the user devised, which adds up a few extra lines of code. but configuring the devised algorithm to a MapReduce job is fairly simple with very less number of steps. The modules are included with different methods for flexibility.

@@

Figure \ref{fig:comp-chart} shows the coding complexities of three image processing computations from Table 1. Single node and the framework show similar coding complexity outperforming the image processing without framework. The reason behind this is Hadoop Image Processing Framework hides the inner details of MapReduce programming which makes an user to think that he is working on the single node but actually the task is processed in hadoop. Without the framework user needs to write a MapReduce job which is cumbersome.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{comp-chart}
	\caption{Comparing coding complexities of laplacian filter, canny edge detection and k-means clustering  in different environments}
	\label{fig:comp-chart}
\end{figure}

Hadoop Image processing framework provides transparency over two levels. The first level of transparency allows user to use the predefined MapReduce modules Downloader, Processor and Extractor for processing the image modules. The second of level of transparency allows the users to write custom map reduce programs as the framework is built in an hierarchal manner.

>>>>

\section{Methodology}
\label{methodology}
The Hadoop Image Processing Framework is intended to provide users
with an accessible, easy-to-use tool for developing large-scale image
processing applications.

@@

\section{Image Processing Algorithms}
\label{algorithms}
To explore the Hadoop Image Processing Framework's capabilities and
performance, we implemented multiple variations of existing
widely-used image processing algorithms:
\begin{itemize}
\item A stand-alone implementation running on a single node with no
  distributed processing capability
\item A generic Hadoop implementation which requires a high level of
  software engineering expertise within the Hadoop framework
\item An implementation that employs the Hadoop Image Processing
  Framework
\end{itemize}

@@

We present execution time and programming complexity results for the
three image processing algorithms described above, demonstrating the
effectiveness of the Hadoop Image Processing Framework.  These
algorithms are not particularly challenging to implement on a single
node, but producing a distributed Hadoop implementation is quite
challenging and requires a great deal of familiarity with Hadoop
concepts and distributed software engineering expertise.  On the other
hand, our framework provides all of the performance and scalability
benefits of the Hadoop system while retaining the simplicity of the
single-node implementation.  In this section we describe the dataset
and hardware and software specifications used in our experiments, and
discuss the results.

@@

\subsection{Characteristics of Image Dataset, Hardware and Software}
We performed image processing computations on a dataset of about 1TB
in size. The dataset was obtained by performing a Flickr search based
on the keyword ``flight''.  The dataset was extracted using the
FlickrRipper, an instance of the Ripper interface provided by our
Hadoop Image Processing Framework.  The framework currently provides
rippers for Flickr and Google Image searches, and more will be
provided in the future.  A Ripper interface extracts image URLs from
search results, and can be implemented for any set of results in any
format.  The experimental dataset is composed of 220000 images at 4.76
MB/image, or about 1 TB.

\subsection{Computer Hardware and Software Characterstics}
We ran the image processing computations on a 6-node distributed
cluster and on a single-node desktop computer. The table below
summarizes the cluster and desktop hardware and software
specifications. The cluster nodes differ in terms of CPU speed and
RAM. We installed Hadoop, Cloudera CDH5 and Java 1.7 on the cluster to
support Java code execution. The desktop computer had a similar
software configuration to the cluster.

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		& \multicolumn{1}{c|}{\textbf{Specs}}                            & \multicolumn{1}{c|}{\textbf{Cluster}}                                                                                                           & \multicolumn{1}{c|}{\textbf{Desktop}}                                                                               \\ \hline
		\multirow{2}{*}{Hardware} & Cluster Nodes                                                  & \begin{tabular}[c]{@{}l@{}}6 data nodes with \\ from 2 to 4 virtual \\ processors and 4 GB\\ RAM and name node \\ with 8 GB RAM\end{tabular} & \begin{tabular}[c]{@{}l@{}}Intel Xeon @ 2Ghz\\ 8 cores, 16GB of RAM \\ and hyperthreading\\ activated\end{tabular} \\ \cline{2-4} 
		& Networking                                                     & 1 Gbit/second                                                                                                                                    &                                                                                                                     \\ \hline
		\multirow{3}{*}{Software} & \begin{tabular}[c]{@{}l@{}}Java Virtual\\ Machine\end{tabular} & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                 \\ \cline{2-4} 
		& Hadoop                                                         & \begin{tabular}[c]{@{}l@{}}Hadoop 2.5\\ Cloudera CDH5\end{tabular}                                                                              &                                                                                                                     \\ \cline{2-4} 
		& \begin{tabular}[c]{@{}l@{}}Operating\\ System\end{tabular}     & CentOS 6.5                                                                                                                                      & CentOS 6.5                                                                                                          \\ \hline
	\end{tabular}
\end{table}

The desktop implementations of the algorithms are Java programs for
processing a small image dataset. As soon as the data set is processed
the program halts. The desktop implementation starts as a single
thread, and images are processed one at a time in succession.  When
the dataset size is low, each program runs smoothly and with excellent
performance.  However, as the size of the data to be processed reaches
the terabyte scale, such a single-process approach fails utterly.
Data on this scale must be processed in distributed fashion using
Hadoop.

Each algorithm is also implemented on a Hadoop cluster without using
the Hadoop Image Processing Framework.  These implementations overcome
the problem of handling large datasets but they suffer from extremely
complex code.  The user needs to write a great deal of code and
understand the technical details of Hadoop. Writing such code,
including custom InputFormats and RecordReaders, is cumbersome and
error-prone. Fig. \ref{fig:module-chart} compares the actual lines of
code a user needs to write in order to run our experimental algorithms
(or any similar image processing application) on a Hadoop cluster.

Finally, we compare the same algorithms implemented on a Hadoop
cluster using the framework.  These implementations overcome the
problems of handling large datasets and managing interoperability
between image datatypes while holding code complexity down to the same
level as the single-process, local implementation.  Users enjoy
automated creation of custom InputFormats and RecordReaders, and can
tune settings for the desired output.  The framework is so transparent
that the user needs only a few lines of code to process a large image
data set. It hides the technical details of MapReduce and runs the
code in highly parallelized fashion.  The code required to implement
each image processing algorithm is no more complex than the same
algorithm implemented on a local, single-threaded system, but the
performance improvement for large datasets is immense.

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		\multicolumn{1}{|c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computa-\\ tional\\ Platform\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computational \\ Elasticity\end{tabular}}}      & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Code \\ Complexity\end{tabular}}}                                                                                                                                                  & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Data \\ Locality\end{tabular}}}                                                                           \\ \hline
		Desktop                                                                                             & \begin{tabular}[c]{@{}l@{}}Low : limited by \\ RAM and CPU of \\ the executing\\ computer\end{tabular} & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing the \\ algorithm\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}All data are \\ on local disk\end{tabular}                                                                                            \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ without \\ using \\ framework\end{tabular}                     & \begin{tabular}[c]{@{}l@{}}High: Nodes can\\  be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Heavy : Complexity \\ lies in writing every \\ module in framework \\ including custom \\ InputFormats and \\ RecordReaders.  User \\ requires high \\ technical expertise \\ with Hadoop.\end{tabular} & \begin{tabular}[c]{@{}l@{}}All data resides\\ in HDFS. Custom\\ InputFormats \\ need to be \\ created to launch\\ computations \\ in appropriate \\ locations\end{tabular} \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ using \\ framework\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}High: Nodes can \\ be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing \\ algorithm and is \\ equivalent to writing \\ on desktop platform\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}All data resides \\ in HDFS. Highly \\ parallelized.\end{tabular}                                                                     \\ \hline
	\end{tabular}
\end{table}

@@

We have explored the space of image processing algorithms, applying
different hardware and software configurations to the various data
sets.  Specifically, we compared:

\begin{itemize}
\item Three image processing algorithms described in Section
  \ref{algorithms}
\item Datasets of varying image counts
\item Configurations of hardware cluster used for computations 
\item Coding complexity (measured in Lines of Code (LOC)
\end{itemize}

The numerical results are shown in Figures \ref{fig:perf-chart}
through \ref{fig:files-chart}.

Figure \ref{fig:perf-chart} illustrates the performance comparison of
Canny edge detection on a single node, on Hadoop using our framework
and Hadoop without using the framework.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{perf-graph}
  \caption{Canny edge detection computation executed on a single node
    and on a Hadoop cluster with and without using framework}
  \label{fig:perf-chart}
\end{figure}

It can be observed that the Hadoop cluster performance is nearly
identical whether the Hadoop Image Processing Framework is used or
not.  The single node performance is satisfactory on small input
sizes; in fact it outperforms Hadoop when the amount of data is very
small owing to the lack of MapReduce overhead.  However, as the amount
of data grows the single node performance becomes increasingly poor.
The same pattern is observed with any image processing algorithm.

Figure \ref{fig:module-chart} compares the coding complexity of the
different image handling and processing tasks described in Section
\ref{methodology}, for the Canny edge detection task. The Downloader
module and Extractor module are simple to use and require few lines of
code to configure. The processor module includes the specific
algorithm devised by the user, which can be significant amount of
code, but using our framework to configure the algorithm for MapReduce
requires almost no additional coding.

@@

Figure \ref{fig:comp-chart} shows the coding complexity of the
processor portion of our three experimental image processing
algorithms.  The naive single-process, single-node algorithm requires
a similar amount of code to using our highly parallel, distributed
framework, while direct implementation of the algorithm using Hadoop
and MapReduce requires a great deal more effort.  The Hadoop Image
Processing Framework hides all of this complexity from the user.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{comp-chart}
  \caption{Comparing coding complexity of Laplacian filter,
    Canny edge detection and $k$-means clustering in different
    environments}
  \label{fig:comp-chart}
\end{figure}

The Hadoop Image processing framework provides transparency over two levels. The first level of transparency allows user to use the predefined MapReduce modules Downloader, Processor and Extractor for processing the image modules. The second of level of transparency allows the users to write custom map reduce programs as the framework is built in an hierarchal manner.
